---
title: "diamonds analysis"
author: "Patrick Ng"
date: "2025-02-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse') ; library('ggplot2')
```

## Introduction

This R Markdown document demonstrates my abilities to use models for data analysis using a data set collected on diamonds.

```{r load data, include = FALSE}
data("diamonds", package = "ggplot2")
names(diamonds)
```

# What is the purpose? 

In this report, we examine the relationship between diamonds' cut, color and clarity.

Let's look at them separately first.

```{r cut, echo = FALSE}
ggplot(diamonds, aes(x = cut)) +
  geom_bar() +
  ggtitle("Distribution of Diamond Cuts") +
  xlab("Cut") +
  ylab("Frequency")
```
```{r color, echo = FALSE}
ggplot(diamonds, aes(x = color)) +
  geom_bar() +
  ggtitle("Distribution of Diamond Colors") +
  xlab("Color") +
  ylab("Frequency")

```

```{r clarity, echo = FALSE}
ggplot(diamonds, aes(x = clarity)) +
  geom_bar() +
  ggtitle("Distribution of Diamond Clarity") +
  xlab("Clarity") +
  ylab("Frequency")

```

Now Graph out them together. I did the same two model accuracy comparsion with anova test.

```{r interactions, echo= FALSE}
ggplot(diamonds, aes(x = cut, y = color, color = clarity)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Interaction of Cut and Color Across Different Clarities") +
  ylab('Cuts') +
  xlab('Colors')

diamonds_model <- lm(price ~ cut * color * clarity, data = diamonds)
summary(diamonds_model)

diamonds_model_simple <- lm(price ~ cut + color + clarity, data = diamonds)
anova(diamonds_model_simple, diamonds_model)

```

The R-squared value is 0.09063, which means about 9.06% of the variation in prices is explained by the model. The p-value associated with this F-statistic is less than 2.2e-16, indicating that the model as a whole is statistically significant.

The p-value in the results is extremely small, which strongly suggests that Model 2 (including the interactions) does a significantly better job at explaining the variations in diamond prices than Model 1.