---
title: "tidy-model"
output: html_document
date: "2025-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse') ; library('modeldata')
``` 

## Introduction

This R Markdown document demonstrates my abilities to use models for data analysis using a data set collected on crickets.

```{r load data, include = FALSE}
data(crickets, package = "modeldata")
names(crickets)
```

# What is that sound? 

In this report, we examine what factors predict a cricket's chirp rate.

```{r summary, echo = FALSE}
summary(crickets)
ggplot(crickets, aes(x = rate)) +
  geom_histogram(bins = 8) + 
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```
* Number of Observations: There are a total of 31 observations in the dataset.
* Number of Species: The dataset covers 2 species of crickets.
* Temperature Range: The temperature ranges from 17.0°C to 30.4°C.
* Mean Rate of Chirping: The average rate of chirping across the dataset is 76.96 chirps per minute.

## Temperature affects chirp rate

```{r temp, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

temp_lm <- lm(rate ~ temp, crickets)

summary.lm(temp_lm)
```

Based on a scatter plot of temperature and chirping and a correlation test, it seems that as temperature increases one degree, the rate of chirping increases about 4.2 chirps per minute.

## Species-specific effects of temperature on chirping

```{r species, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')
  
  
species_lm <- lm(rate ~ temp + species, crickets)

summary.lm(species_lm)
```

The second model, which looks at different cricket species, explains about 99% of why chirp rates vary, showing it's a very accurate predictor. This means if we know the species, we can almost perfectly guess how often a cricket chirps. It's a more reliable tool for predictions compared to the first model based on temperature, which explains about 92% of the variations.

## Interactions

```{r species historgram, echo = FALSE}
ggplot(crickets, aes(x = rate, fill = species)) +
  geom_histogram(position = 'identity', alpha = 0.7, bins = 8) + 
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```

```{r interactions, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')
  
species_x_temp_lm <- lm(rate ~ temp + species + temp:species, crickets) 
summary.lm(species_x_temp_lm)

anova(species_lm, species_x_temp_lm)
```